Synthesizing Quotient Lenses

### 2017-10-15 papers/QuotientLensSynth/main.tex
add eval rough draft

@@

For example, in the \textsc{Bib}\TeX\ to EndNote problem, the programmer wishes
to synthesize a QRE lens $q$ which maps the QRE
$$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
\re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$ 
to the QRE
$$\perm{\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},
\re{YEAR'}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$
(resp. $\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},$ and
$\re{YEAR'}$) are regular expressions describing the respective
\textsc{Bib}\TeX\; (resp. EndNote\bcp{Don't like this ``resp.''
  construction---saves space, but not reader energy}) fields of the same
name, and 
$\re{WHITESPACE}$ is a regular expression describing whitespace
characters. Morever, the synthesized lens should map the equivalence
class of this \textsc{Bib}\TeX{} record
\begin{verbatim}
@Book {conway,
  Author = {Conway, J. H.},
  Title = {Regular Algebra and Finite Machines},
  Publisher = {Printed in GB by William Clowes & Sons Ltd},
  Year = {1971}
}
\end{verbatim}
\noindent to the equivalence class of this EndNote entry:
\begin{verbatim}
%0 Book
%T Regular Algebra and Finite Machines
%A Conway, J. H.
%D 1971
%I Printed in GB by William Clowes & Sons Ltd
%F conway
\end{verbatim}

@@

>>>>

% Evaluation AST Metrics
\newcommand{\QRESize}{\textbf{QS}}
\newcommand{\CanonizerAndSpecSize}{\textbf{TCS}}
\newcommand{\LensAndSpecSize}{\textbf{TLS}}

% Evaluation Runtime Metrics
\newcommand{\SystemOnBenchmarks}{\textbf{MM}}
\newcommand{\SystemOnOptician}{\textbf{MO}}
\newcommand{\OpticianRuntime}{\textbf{OO}}

\newcommand{\Name}{Optometrist}

@@

For example, in the \textsc{Bib}\TeX\ to EndNote problem, the programmer wishes
to synthesize a QRE lens $q$ which maps the QRE
$$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
  \re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$ 
to the QRE
$$\perm{\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},
  \re{YEAR'}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$
(resp. $\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},$ and
$\re{YEAR'}$) are regular expressions describing the respective
\textsc{Bib}\TeX\; (resp. EndNote\bcp{Don't like this ``resp.''
  construction---saves space, but not reader energy}) fields of the same
name, and 
$\re{WHITESPACE}$ is a regular expression describing whitespace
characters. Morever, the synthesized lens should map the equivalence
class of this \textsc{Bib}\TeX{} record
\begin{verbatim}
@Book {conway,
  Author = {Conway, J. H.},
  Title = {Regular Algebra and Finite Machines},
  Publisher = {Printed in GB by William Clowes & Sons Ltd},
  Year = {1971}
}
\end{verbatim}
\noindent to the equivalence class of this EndNote entry:
\begin{verbatim}
% 0 Book
% T Regular Algebra and Finite Machines
% A Conway, J. H.
% D 1971
% I Printed in GB by William Clowes & Sons Ltd
% F conway
\end{verbatim}

@@

We have integrated QREs and quotient lens synthesis into Boomerang with an
additional \afm{loc} lines of OCaml code.

All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS Sierra.

Our evaluations serve to answer the following questions:
\begin{enumerate}
\item Does synthesizing quotient lenses from QREs permit for an easier
  development process than existing approaches.

\item Are QREs sufficiently expressive that more lenses can be synthesized
  without modifications to the format.

\item Is our synthesis algorithm competitive with existing lens synthesis
  tools.
\end{enumerate}

\subsection{Benchmark Suite Construction} 
Our benchmarks were constructed from two sources:
\begin{enumerate}
\item \afm{39} of our benchmarks were derived from the Optician system.  These
  benchmarks were a combination of custom benchmarks, benchmarks derived from
  FlashFill, and benchmarks derived from Augeas.  Many of the formats mapped
  between had to be modified to work with the bijectivity constraints
  Optician required.  Commonly, one format had whitespace where the other did
  not allow for whitespace.  To address the bijectivity issues surrounding this,
  we added whitespace to the format that typically did not have whitespace.  We
  were able to remove these alterations in adopting these benchmarks to our
  system.

\item \afm{10} of our benchmarks were derived from data.gov.  Data.gov contains
  many transformations between formats.  In addition to the formats previously
  adopted for Optician, we synthesized transformers between these formats.
\end{enumerate}

\subsection{Easing the Development Process}

To test if our tool eases the development process, we compare the following
number of AST nodes.
%
\begin{itemize}
\item[\QRESize{}] \QRESize{} is the number of AST nodes in the QRE
  specifications involved in QRE lens synthesis task.
\item[\CanonizerAndSpecSize{}] \CanonizerAndSpecSize{} is the number of
  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
  for each QRE used in QRE lens synthesis task.
\item[\LensAndSpecSize{}] \LensAndSpecSize{} is the number of AST nodes
  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
\end{itemize}

The value of \QRESize{} represents the difficulty in writing our benchmarks.

The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
benchmarks without QREs.  This is only an approximation, as both $W(q)$ and
$\canonizer(q)$ are automatically generated from the QRE.  In our
experience, these automatically generated regular expressions and canonizers are
the ones we would naturally write.

The value of \LensAndSpecSize{} represents the difficulty in writing our benchmarks
with neither QREs nor synthesis.  This is also an approximation, as $W(q)$ and
$\canonizer(q)$ and the bijective lens are all automatically generated.
We have found that the generated bijective lens is fairly similar in size to the
one we would manually write, though it is typically less modular.

Figure~\ref{?} shows the number of benchmarks written with at most a given
number of AST nodes.  We found that \QRESize{} was consistently smaller than
\CanonizerAndSpecSize{} and \LensAndSpecSize{}, demonstrating that introducing
QREs saves programmers programming effort compared to both Optician and basic
Boomerang.

\subsection{Increase of Expressivity}

Optician was only able to synthesize fully bijective data transformations.
\Name loosens that restriction, and is able to synthesize data which is
bijective when put in canonical form.
To determine the increase of expressivity, we analyzed:
\begin{enumerate}
\item How many data formats in Optician's benchmark suite required modifications
  to be bijective.
\item How many data formats in Optician's benchmark suite required modifications
  to be bijective when put in canonical form.
\end{enumerate}

We found that \afm{35} of Opticians 39 benchmarks required modifications to be
bijective when put in canonical form. We found that \afm{25} of Optician's 39
benchmarks required modifications to be in bijective correspondence when in
canonical form.  Furthermore, we found that \afm{all} of the data.gov examples
required no modifications to be in canonical form.  While we did not put the
data.gov examples into a form suitable for Optician, we found only \afm{1} of
the \afm{5} data.gov examples did not require canonization, so making suitable
for Optician would have required a modification of \afm{4} of the data.gov
examples.

\subsection{Competitive Performance}

While we expanded on the Optician system, we found it important to validate
performance did not degrade in this expansion.  To validate this, we tested:
\begin{itemize}
\item[\OpticianRuntime{}] Optician runs on its own benchmarks.
\item[\SystemOnOptician{}] \Name{} runs on Optician's benchmarks.
\item[\SystemOnBenchmarks{}] \Name{} runs on its own benchmarks.
\end{itemize}

We summarize the results of these experiments in Figure~\ref{?}.  \Name{} was
able to synthesize all of Optician's benchmarks at a competitive speed with
Optician.  Furthermore, adopting these tasks to use QREs actually sped up
tasks!  After canonization, the regular expressions sent to synthesis are
\emph{smaller} than before canonization - making the bijective lens synthesis
portion of these tasks easier.
